Flowchart of Current Algorithm

### 2012-04-18 docs/flowchart/flowchart.tex
latex fixes
some notes

Here we summarize the equations solved by \maestro.  We refer the reader
to papers~I through IV and the multilevel paper for the derivation
and motivation of the equation set.

@@

\section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
Here is the current description of the algorithm, based from the
description in the multilevel paper.  The initialization has also been
included with more detail than was given in paper III.  The main
driver for a single step of the algorithm is {\tt advance.f90}---refer
to that code to see the sequence of functions called to implement each
ste.

@@

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$.  We use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  The perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  is used to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

%-----------------------------------------------------------------------------
% graphical flowcharts
%-----------------------------------------------------------------------------
\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\flowfigpath/flowchart}
\caption[Graphical flowchart of MAESTRO]
  {\label{Fig:flowchart} A flowchart of the algorithm.  The
  thermodynamic state variables, base state variables, and local velocity are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  The predictor-corrector steps are
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The diffusion steps (4a and 8a) are optional, depending on
  {\tt use\_thermal\_diffusion}.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]                                               
\centering
\includegraphics[scale=0.75]{\flowfigpath/flowchart_4_8}
\caption[Graphical flowchart of the density and enthalpy update steps]
{\label{Fig:flowchart48} A flowchart for {\bf Steps 4} and {\bf 8}.
  The thermodynamic state variables and base state variables are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  Note, for {\bf Step 4}, the updated
  quantities should also have a $\star$ superscript, e.g., {\bf Step
    8I} defines $T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$
  .}
\end{figure}

>>>>

Here we summarize the equations solved by \maestro.  We refer the reader
to papers~I through IV and the multilevel paper for the derivation
and motivation of the equation set.  (Note: this `traditional' algorithm
uses Strang-splitting for the reactions.  An alternate implementation, using
spectral deferred corrections is outlined in Chapter~\ref{ch:sdc}.)

@@

\section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
Here is the current description of the algorithm, based from the
description in the multilevel paper.  The initialization has also been
included with more detail than was given in paper III.  The main
driver for a single step of the algorithm is {\tt advance.f90}---refer
to that code to see the sequence of functions called to implement each
step.

@@

reach\_state

enforce\_hse

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$.

  There are a variety of choices of quantities to predict to the
  edges (controlled by {\tt species\_pred\_type}---see Chapter~\ref{ch:pert}).
  By default, we use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
   of quantities that we can predict to the interfaces here (controlled
   by {\tt enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
   the default case, we use the perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

%-----------------------------------------------------------------------------
% graphical flowcharts
%-----------------------------------------------------------------------------
\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\flowfigpath/flowchart}
\caption[Graphical flowchart of \maestro]
  {\label{Fig:flowchart} A flowchart of the algorithm.  The
  thermodynamic state variables, base state variables, and local velocity are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  The predictor-corrector steps are
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The diffusion steps (4a and 8a) are optional, depending on
  {\tt use\_thermal\_diffusion}.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]                                               
\centering
\includegraphics[scale=0.75]{\flowfigpath/flowchart_4_8}
\caption[Graphical flowchart of the density and enthalpy update steps]
{\label{Fig:flowchart48} A flowchart for {\bf Steps 4} and {\bf 8}.
  The thermodynamic state variables and base state variables are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  Note, for {\bf Step 4}, the updated
  quantities should also have a $\star$ superscript, e.g., {\bf Step
    8I} defines $T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$
  .}
\end{figure}
